Comments on Blair and Sambanis, 2020, \"Forecasting Civil Wars: Theory and Structure in an Age of 'Big Data' and Machine Learning\", *JCR*

### 2020-06-06 paper.Rmd
add text about B&S table 4 to paper

THIS SECTION IS EMPTY

@@

tab4 %>%
  knitr::kable(
    caption = "Replication of B\\&S Table 4: 2016 Confusion Matrices for Six-month Escalation Model.",
    booktabs = TRUE
  ) %>%
  kableExtra::collapse_rows(1:2, row_group_label_position = "stack")
```

>>>>

B&S report confusion matrices for their forecasts for the first half of 2016 (2016-H1) in their Table 4. To create the confusion matrices, B&S treat the 30 highest ranked predictions as positive predictions ("1s") and the rest as negative predictions ("0s"). We replicate their Table 4 in the top portion of \ref{tab:table4}. There are two confusion matrices for slightly different codings of outcomes in 2016-H1, under the corresponding "Assuming Persistence" and "Assuming Change" headings. We should not that the "Assuming Persistence" corresponds to the values in the replication data; the small variations for the "Assuming Change" version are hand-coded in the replication code file that generates the confusion matrices.

We can see that the original table presents 15 or 16 positive cases for 2016-H1, depending on teh DV coding variation. This corresponds to a positive rate of around 9.5% for the first half of 2016 data. In contrast, the corresponding 6-month version of the data from 2001 to 2015, with 30 half-years, has in total 20 civil war onset events, for a much lower positive rate of around 0.5%. Since the positive event rate in the confusion matrices far exceeds the rate of observed civil war onsets in both the training and test data, we suspected that the forecasts were erroneously assessed using civil war incidence, not onset. As we showed above, by examining the replication code we were able to verify that the forecasts were scored using civil war incidence, i.e. including ongoing civil wars, rather than civil war onset years only.

tab4 %>%
  knitr::kable(
    caption = "Replication of B\\&S Table 4: 2016 Confusion Matrices for Six-month Escalation Model.",
    label = "table4",
    booktabs = TRUE
  ) %>%
  kableExtra::collapse_rows(1:2, row_group_label_position = "stack")
```

The correct confusion matrices when using observed onset (or the lack of it) are shown in the second part of Table \ref{tab:table4}. In the default "Assuming Persistence" coding, there are no civil war onsets in the data for 2016-H1. Thus the recall values is undefined, while the precision is 0/30 = 0, compared to recall and precision values of 13/15 = 0.87 and 13/30 = 0.43 with the incorrect scoring. The alternative coding ("Assuming Change") produces 2 civil war onsets. Recall is 1 compared to 14/16 = 0.88 before, and precision is 2/30 = 0.07 instead of 14/30 = 0.47.

Another, minor issue or rather coding error, is related to using a lead version of the DV. With the lead version of the DV, "incidence_civil_ns_plus1", which is what the models are predicting, the predicted value for 2016-H1 actually indicates the risk of civil war onset in 2016-H2. In the Table 4 script referenced above, the 2016-H1 predictions (for 2016-H2) are assessed using the raw DV, "incidence_civil_ns", not the lead version. Essentially, the forecasts for 2016-H2 are assessed using observed outcomes for 2016-H1. In this case it doesn't make a difference since both the raw DV and lead version for 2016-H1 do not have any positive values.

@@
